---
title: "Peer_assignment_1"
author: "GabriellaJacoel"
date: "16 luglio 2017"
output: html_document
---

## Daily Activity Analysis

First of all, I import the data through function 'read.csv' and I convert the data.frame into data.table, for simplicity.

```{r}
library(data.table)
setwd("C:/Users/GJacoel/Desktop/Coursera/Reproducible Research/Week2/")
activity<-read.csv("activity.csv")
activity<-as.data.table(activity)
daily_activity<-activity[,list(daily_step=sum(steps,na.rm = T)),by=date]

```
Then, I aggregate data, summing up the number of steps for each day.
```{r}
daily_activity<-activity[,list(daily_step=sum(steps,na.rm = T)),by=date]
```

In order to study daily steps distribution, we plot the histogram of steps.

```{r pressure, echo=FALSE}
hist(daily_activity$daily_step,col = "firebrick2",xlab ="Number of daily steps",
     main ="Distribution of daily steps")
```

```{r computetime,echo=TRUE}
daily_mean<-mean(daily_activity$daily_step)
daily_median<-median(daily_activity$daily_step)
```

Mean of distribution is `r daily_mean`.
Median of distribution is `r daily_median`

## Average daily activity pattern
I compute the average number of steps for each 5 minute interval, evaluated over all the days at our disposal:
```{r}
av_daily_activity<-activity[,list(av_step=mean(steps,na.rm = T)),by=interval]

```
Then, we plot the computed points on a line-plot, that shows the average pattern of daily activity:
```{r}
plot(y=av_daily_activity$av_step,x = av_daily_activity$interval,type = "l",col="darkcyan",xlab = "5 MINUTE INTERVAL",ylab = "AVERAGE NUMBER OF STEPS",main = "Average number of steps during the day")
```

## Missing values analysis
```{r}
x<-sapply(activity,function(x) sum(is.na(x)))
```
Total number of missing values in the dataset is `r x[1]`. We replace all the missing values concentrated on "Steps" column, with the average number of steps calculated in the above analysis:

```{r}
no_na_activity<-merge(x=activity,y=av_daily_activity, by=c("interval"),all.x = T)
no_na_activity$steps<-as.numeric(no_na_activity$steps)
no_na_activity[is.na(no_na_activity$steps),]$steps<-no_na_activity[is.na(no_na_activity$steps),]$av_step
```
Now we recompute the daily analysis, using the new imputed values:

```{r}
no_na_daily_activity<-no_na_activity[,list(daily_step=sum(steps,na.rm = T)),by=date]
```

In order to study daily steps distribution, we plot the histogram of steps.

```{r}
hist (no_na_daily_activity$daily_step,col = "orange",xlab ="Number of daily steps", main ="Distribution of daily steps (no missing values)")
```

```{r,echo=TRUE}
no_na_daily_mean<-mean(no_na_daily_activity$daily_step)
no_na_daily_median<-median(no_na_daily_activity$daily_step)
```

Mean of distribution is `r no_na_daily_mean`.
Median of distribution is `r no_na_daily_median`

##Analysis weekend/weekdays activity
We classify each date as "Weekend" or "Weekday" to perform the analysis:

```{r,echo=TRUE}
no_na_activity$date<-as.Date(no_na_activity$date)
no_na_activity$weekdays<-weekdays(no_na_activity$date)
no_na_activity$weekend<-"Weekday"
no_na_activity[no_na_activity$weekdays=="sabato"|no_na_activity$weekdays=="domenica",]$weekend<-"Weekend"
no_na_activity$weekdays<-NULL

no_na_av_daily_activity<-no_na_activity[,list(av_step=mean(steps,na.rm = T)),by=c("interval","weekend")]
weekend_activity<-no_na_av_daily_activity[weekend=="Weekend"]
weekday_activity<-no_na_av_daily_activity[weekend=="Weekday"]
```

Now we plot the results:

```{r}
par(mfrow=c(1,2))
plot(y=weekend_activity$av_step,x = weekend_activity$interval,type = "l",col="blue",xlab = "5 MINUTE INTERVAL",ylab = "AVERAGE NUMBER OF STEPS", main="Weekend activity")
plot(y=weekday_activity$av_step,x = weekday_activity$interval,type = "l",col="coral",xlab = "5 MINUTE INTERVAL",ylab = "AVERAGE NUMBER OF STEPS", main="Weekday activity")
```

